
# Task 3 (Sales Prediction, Dealership Performance Analysis, and Ensemble Learning)
Person 3: Sales Prediction, Dealership Performance Analysis, and Ensemble Learning
Sales Prediction Model

Develop a predictive model for car sales.
Dealership Performance Analysis

Evaluate dealership performance across regions.
Ensemble Learning

Implement ensemble learning methods to improve the accuracy and robustness of the sales prediction model by combining multiple predictive models.

Reading Dataset


```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(stringr)
library(fastDummies)
library(lubridate)
library(caret)
cars_data <- read.csv("car_data.csv", head = TRUE)
cars_data <- cars_data %>%
    mutate(
      Date = mdy(Date), # Convert Date to Date format assuming m/d/y format
        Engine = gsub("Â", "", Engine), # Clean Engine column
        Price = as.numeric(Price) # Ensure numeric
)

cars_data$Date <- as.Date(cars_data$Date)
cars_data$month <- month(cars_data$Date)
cars_data$season <- quarter(cars_data$Date, with_year = FALSE)
```

```{r}
colnames(cars_data)
```

```{r}
cars_data["Dealer_Region"]
```


```{r}
cars_data <- cars_data %>%
  mutate(
    Gender = as.numeric(factor(Gender, levels = c("Male", "Female"), labels = c(1, 0))),
    Dealer_Name = as.numeric(factor(Dealer_Name)),
    Company = as.numeric(factor(Company)),
    Model = as.numeric(factor(Model)),
    Transmission = as.numeric(factor(Transmission)),
    Color = as.numeric(factor(Color)),
    Body_Style = as.numeric(factor(Body_Style)),
    Dealer_Region = as.numeric(factor(Dealer_Region))
  )
```


```{r }
set.seed(6659)

filter_data <- select(cars_data, -Car_id, -Date, -Customer_Name, -Engine, -Phone, -Dealer_No, -Dealer_Region)

train_indices <- sample(1:nrow(filter_data), 0.7 * nrow(filter_data)) 
train_data <- filter_data[train_indices, ]
test_data <- filter_data[-train_indices, ]
model <- lm(Price ~ ., data = train_data)

anova(model)
```
```{r }
colnames(filter_data)

anova(model)
```


```{r }
filter_data <- select(cars_data, -Car_id, -Date, -Customer_Name, -Engine, -Phone, -Dealer_No, -Dealer_Region, -month)

train_data <- filter_data[train_indices, ]

test_data <- filter_data[-train_indices, ]
model <- lm(Price ~ ., data = train_data)

anova(model)
```



```{r }
filter_data <- select(cars_data, -Car_id, -Date, -Customer_Name, -Engine, -Phone, -Dealer_No, -Dealer_Region, -month, -Gender)

train_data <- filter_data[train_indices, ]

test_data <- filter_data[-train_indices, ]
model <- lm(Price ~ ., data = train_data)

anova(model)
```


```{r }
filter_data <- select(cars_data, -Car_id, -Date, -Customer_Name, -Engine, -Phone, -Dealer_No, -Dealer_Region, -month, -season, -Gender)

train_data <- filter_data[train_indices, ]

test_data <- filter_data[-train_indices, ]
model <- lm(Price ~ ., data = train_data)

anova(model)
```

```{r }
filter_data <- select(cars_data, -Car_id, -Date, -Customer_Name, -Engine, -Phone, -Dealer_No, -Dealer_Region, -month, -season, -Gender, -Color)

train_data <- filter_data[train_indices, ]

test_data <- filter_data[-train_indices, ]
model <- lm(Price ~ ., data = train_data)

anova(model)
```


```{r }
filter_data <- select(cars_data, -Car_id, -Date, -Customer_Name, -Engine, -Phone, -Dealer_No, -Dealer_Region, -month, -season, -Gender, -Dealer_Name, -Color)

train_data <- filter_data[train_indices, ]

test_data <- filter_data[-train_indices, ]
model <- lm(Price ~ ., data = train_data)

anova(model)
```

```{r }
filter_data <- select(cars_data, -Car_id, -Date, -Customer_Name, -Engine, -Phone, -Dealer_No, -Dealer_Region, -month, -season, -Gender, -Dealer_Name, -Color)

train_data <- filter_data[train_indices, ]

test_data <- filter_data[-train_indices, ]
model <- lm(Price ~ ., data = train_data)

anova(model)
```

```{r }
filter_data <- select(cars_data, -Car_id, -Date, -Customer_Name, -Engine, -Phone, -Dealer_No, -Dealer_Region, -month, -season, -Gender, -Dealer_Name, -Color, -Body_Style)

train_data <- filter_data[train_indices, ]

test_data <- filter_data[-train_indices, ]
model <- lm(Price ~ ., data = train_data)

anova(model)
```

```{r }
max(filter_data[,c("Price")])


```

```{r }
predictions <- predict(model, newdata = train_data)
threshold <- 500
accuracy <- mean(abs(predictions - train_data$Price) <= threshold)
print(paste("Accuracy (within $500):", accuracy))


```



```{r}
set.seed(6659) 
filter_data <- select(cars_data, -Car_id, -Date, -Customer_Name, -Engine, -Dealer_No, -Phone)
train_index <- sample(1:nrow(filter_data), 0.7 * nrow(filter_data)) 
train_data <- filter_data[train_index, ]
test_data <- filter_data[-train_index, ]
```

```{r }
head(filter_data)
```

```{r}
library(randomForest)
rf_model <- randomForest(Price ~ ., data = train_data, ntree = 500, mtry = 30, nodesize = 15)
predictions <- predict(rf_model, newdata = test_data)
threshold <- 500
accuracy <- mean(abs(predictions - test_data$Price) <= threshold)
print(paste("Accuracy (within $500):", accuracy))
```

```{r}
threshold <- mean(abs(predictions - test_data$Price))
accuracy <- mean(abs(predictions - test_data$Price) <= threshold)
print(paste("Accuracy (within $500):", accuracy))
```

```{r}
mean(abs(predictions - test_data$Price))

```


```{r}
threshold <- 500
accuracy <- mean(abs(predictions - test_data$Price) <= threshold)
print(paste("Accuracy (within $500):", accuracy))
```


```{r}
library(dplyr)  # For data manipulation
library(ggplot2)  # For visualization

customer_segmentation <- cars_data %>%
  group_by(Gender, Company, Model) %>%
  summarise(total_sales = n()) %>%
  arrange(desc(total_sales))

```


```{r}
colnames(cars_data)
```

```{r}
customer_segmentation
```

```{r}
top_selling_models <- cars_data %>%
  group_by(Company, Model) %>%
  summarise(total_sales = n()) %>%
  arrange(desc(total_sales)) %>%
  top_n(10)

```

```{r}
library(ggplot2)
library(dplyr)

# Assuming 'car_data' is your dataset

# Group data by Dealer_Region and Model, then calculate total sales for each combination
sales_by_region <- cars_data %>%
  group_by(Dealer_Region, Company) %>%
  summarise(total_sales = n()) %>%
  arrange(Dealer_Region, desc(total_sales)) %>%
  group_by(Dealer_Region) %>%
  top_n(3, total_sales)  # Select the top selling car model for each region

# Plot the most sold cars by Dealer_Region
ggplot(sales_by_region, aes(x = Dealer_Region, y = total_sales, fill = Company)) +
  geom_bar(stat = "identity") +
  labs(title = "Most Sold Cars by Dealer Region", x = "Dealer Region", y = "Total Sales") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_viridis_d()  # Change the color scale if desired
```

```{r}
average_price_by_region <- cars_data %>%
  group_by(Dealer_Region) %>%
  summarise(avg_price = mean(Price, na.rm = TRUE))

# Define a color palette
region_colors <- c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd", "#8c564b", "#e377c2", "#7f7f7f", "#bcbd22", "#17becf")

# Plot the average price by Dealer_Region with specified colors
ggplot(average_price_by_region, aes(x = Dealer_Region, y = avg_price, fill = Dealer_Region)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Price by Dealer Region", x = "Dealer Region", y = "Average Price") +
  scale_fill_manual(values = region_colors) +  # Specify the color palette
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r}
ggplot(cars_data, aes(x = Dealer_Region, fill = Body_Style)) +
  geom_bar(position = "stack") +
  labs(title = "Body Style Distribution by Dealer Region", x = "Dealer Region", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_viridis_d()
```
