---
title: "Project Prediction"
author: "Mauricio Gómez Macedo"
date: "April 7, 2024"
output: html_document
---

Person 3: Sales Prediction, Dealership Performance Analysis, and Ensemble Learning
Sales Prediction Model

Develop a predictive model for car sales.
Dealership Performance Analysis

Evaluate dealership performance across regions.
Ensemble Learning

Implement ensemble learning methods to improve the accuracy and robustness of the sales prediction model by combining multiple predictive models.

Reading Dataset

```{r }
cars_df <- read.csv("cars_data_encoded.csv")

```

```{r }
head(cars_df)
```

Selecting variables

```{r }
df <- cars_df[,c("Price","Gender", "Annual_Income", "Transmission_Auto",	"Transmission_Manual",	"Color_Black",	"Color_Pale_White",	"Color_Red",	"Body_Style_Hardtop",	"Body_Style_Hatchback",	"Body_Style_Passenger",	"Body_Style_Sedan",	"Body_Style_SUV",
"Engine_Double_Overhead_Camshaft",
"Engine_Overhead_Camshaft",	"Dealer_Region_Aurora",	"Dealer_Region_Austin",	"Dealer_Region_Greenville",	"Dealer_Region_Janesville",	"Dealer_Region_Middletown",	"Dealer_Region_Pasco",	"Dealer_Region_Scottsdale",	"Car_Age"
)]
```


```{r }
head(df)
```

Splitting Data. 80% Train, 20% Test

```{r }
set.seed(6659)
total_rows <- nrow(df)

# Number of rows for training (80%)
train_size <- round(0.8 * total_rows)

# Indices of rows to be selected for training
train_indices <- sample(1:total_rows, train_size, replace = FALSE)

# Create training dataset
train_data <- df[train_indices, ]

# Create testing dataset
test_data <- df[-train_indices, ]
```


Linear Regression


```{r }
model <- lm(Price ~ ., data = train_data)

```

```{r }
anova(model)
```

Removing variables (1 by 1)

Removing Dealer_Region_Middletown

```{r }
df <- cars_df[,c("Price","Gender", "Annual_Income", "Transmission_Auto",	"Color_Black",	"Color_Pale_White",	"Body_Style_Hardtop",	"Body_Style_Hatchback",	"Body_Style_Passenger",	"Body_Style_Sedan",	"Dealer_Region_Aurora",	"Dealer_Region_Austin",	"Dealer_Region_Greenville",	"Dealer_Region_Janesville",	"Car_Age"
)]

train_size <- round(0.8 * total_rows)
train_indices <- sample(1:total_rows, train_size, replace = FALSE)
train_data <- df[train_indices, ]
test_data <- df[-train_indices, ]

model <- lm(Price ~ ., data = train_data)

anova(model)
```

Removing Dalear_Region_Janesville

```{r }
df <- cars_df[,c("Price","Gender", "Annual_Income", "Transmission_Auto",	"Color_Black",	"Color_Pale_White",	"Body_Style_Hardtop",	"Body_Style_Hatchback",	"Body_Style_Passenger",	"Body_Style_Sedan",	"Dealer_Region_Aurora",	"Dealer_Region_Austin",	"Dealer_Region_Greenville",	"Car_Age"
)]

train_size <- round(0.8 * total_rows)
train_indices <- sample(1:total_rows, train_size, replace = FALSE)
train_data <- df[train_indices, ]
test_data <- df[-train_indices, ]

model <- lm(Price ~ ., data = train_data)

anova(model)
```

Removing Gender

```{r }
df <- cars_df[,c("Price", "Annual_Income", "Transmission_Auto",	"Color_Black",	"Color_Pale_White",	"Body_Style_Hardtop",	"Body_Style_Hatchback",	"Body_Style_Passenger",	"Body_Style_Sedan",	"Dealer_Region_Aurora",	"Dealer_Region_Austin",	"Dealer_Region_Greenville",	"Car_Age"
)]

train_size <- round(0.8 * total_rows)
train_indices <- sample(1:total_rows, train_size, replace = FALSE)
train_data <- df[train_indices, ]
test_data <- df[-train_indices, ]

model <- lm(Price ~ ., data = train_data)

anova(model)
```

Removing Car_Age

```{r }
df <- cars_df[,c("Price", "Annual_Income", "Transmission_Auto",	"Color_Black",	"Color_Pale_White",	"Body_Style_Hardtop",	"Body_Style_Hatchback",	"Body_Style_Passenger",	"Body_Style_Sedan",	"Dealer_Region_Aurora",	"Dealer_Region_Austin",	"Dealer_Region_Greenville"
)]

train_size <- round(0.8 * total_rows)
train_indices <- sample(1:total_rows, train_size, replace = FALSE)
train_data <- df[train_indices, ]
test_data <- df[-train_indices, ]

model <- lm(Price ~ ., data = train_data)

anova(model)
```

Removing Dealear_Region_Aurora

```{r }
df <- cars_df[,c("Price", "Annual_Income", "Transmission_Auto",	"Color_Black",	"Color_Pale_White",	"Body_Style_Hardtop",	"Body_Style_Hatchback",	"Body_Style_Passenger",	"Body_Style_Sedan",	"Dealer_Region_Austin",	"Dealer_Region_Greenville"
)]

train_size <- round(0.8 * total_rows)
train_indices <- sample(1:total_rows, train_size, replace = FALSE)
train_data <- df[train_indices, ]
test_data <- df[-train_indices, ]

model <- lm(Price ~ ., data = train_data)

anova(model)
```

Removing Dealear_Region_Greenville

```{r }
df <- cars_df[,c("Price", "Annual_Income", "Transmission_Auto",	"Color_Black",	"Color_Pale_White",	"Body_Style_Hardtop",	"Body_Style_Hatchback",	"Body_Style_Passenger",	"Body_Style_Sedan",	"Dealer_Region_Austin"
)]

train_size <- round(0.8 * total_rows)
train_indices <- sample(1:total_rows, train_size, replace = FALSE)
train_data <- df[train_indices, ]
test_data <- df[-train_indices, ]

model <- lm(Price ~ ., data = train_data)

anova(model)
```

Removing Dealear_Region_Austin

```{r }
df <- cars_df[,c("Price", "Annual_Income", "Transmission_Auto",	"Color_Black",	"Color_Pale_White",	"Body_Style_Hardtop",	"Body_Style_Hatchback",	"Body_Style_Passenger",	"Body_Style_Sedan"
)]

train_size <- round(0.8 * total_rows)
train_indices <- sample(1:total_rows, train_size, replace = FALSE)
train_data <- df[train_indices, ]
test_data <- df[-train_indices, ]

model <- lm(Price ~ ., data = train_data)

anova(model)
```

Removing Transmission_Auto

```{r }
df <- cars_df[,c("Price", "Annual_Income",	"Color_Black",	"Color_Pale_White",	"Body_Style_Hardtop",	"Body_Style_Hatchback",	"Body_Style_Passenger",	"Body_Style_Sedan"
)]

train_size <- round(0.8 * total_rows)
train_indices <- sample(1:total_rows, train_size, replace = FALSE)
df_scale <- as.data.frame(scale(df))
train_data <- df_scale[train_indices, ]
test_data <- df_scale[-train_indices, ]

model <- lm(Price ~ ., data = train_data)

anova(model)
```


```{r }
predictions <- predict(model, newdata = test_data)
```

```{r }
library(Metrics)

rmse_value <- rmse(test_data$Price, predictions)

```


```{r }
summary(model)$r.squared

```


```{r }
library(randomForest)

df <- cars_df[,c("Price","Gender", "Annual_Income", "Transmission_Auto",	"Transmission_Manual",	"Color_Black",	"Color_Pale_White",	"Color_Red",	"Body_Style_Hardtop",	"Body_Style_Hatchback",	"Body_Style_Passenger",	"Body_Style_Sedan",	"Body_Style_SUV",
"Engine_Double_Overhead_Camshaft",
"Engine_Overhead_Camshaft",	"Dealer_Region_Aurora",	"Dealer_Region_Austin",	"Dealer_Region_Greenville",	"Dealer_Region_Janesville",	"Dealer_Region_Middletown",	"Dealer_Region_Pasco",	"Dealer_Region_Scottsdale",	"Car_Age"
)]

train_data <- df[train_indices, ]
test_data <- df[-train_indices, ]
rf_model <- randomForest(Price ~ ., data = train_data)




```

```{r }
predictions <- predict(rf_model, newdata = test_data)

rmse_value <- sqrt(mean((test_data$Price - predictions)^2))

# Calculate R-squared (R2)
mean_actual <- mean(test_data$Price)
ss_total <- sum((test_data$Price - mean_actual)^2)
ss_residual <- sum((test_data$Price - predictions)^2)
r_squared_value <- 1 - (ss_residual / ss_total)

# Print RMSE and R-squared (R2)
print(paste("RMSE:", rmse_value))
print(paste("R-squared (R2):", r_squared_value))

```