---
title: "Project"
author: "Chen Chen"
date: "2024-04-06"
output: html_document
---

# Introduction
The dataset we used is from Kaggle, which is a car sale report among same certain area of United States. There are 16 columns, including car id, purchase date, customer name, gender, annual income, dealer name, car company, car model, engine, transmission, color, price, dealer number, body style, phone and dealer region. 

```{r, warning = FALSE, include = FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(stringr)
library(fastDummies)
library(lubridate)
library(caret)
library(shiny)
```

```{r}
# load dataset
cars_data <- read.csv("car_data.csv", head = TRUE)
head(cars_data)
```

```{r}
cars_data <- cars_data %>%
    mutate(
      Date = mdy(Date), # Convert Date to Date format assuming m/d/y format
        Engine = gsub("Â", "", Engine), # Clean Engine column
        Price = as.numeric(Price....), # Ensure numeric
    )
cars_data$Date <- as.Date(cars_data$Date)
cars_data$month <- month(cars_data$Date)
cars_data$season <- quarter(cars_data$Date, with_year = FALSE)
```

# Interactive Dashboard
```{r}
# Lookup history car price via different dealer and car type
ui_his_price <- fluidPage(
  titlePanel("History Car Price Lookup"),
  sidebarLayout(
    sidebarPanel(
      selectInput("dealer_name", "Select Dealer Name:", choices = unique(cars_data$Dealer_Name)),
      selectInput("company", "Select Car Company:", choices = unique(cars_data$Company)),
      selectInput("model", "Select Car Model:", choices = unique(cars_data$Model)),
      selectInput("transmission", "Select Transmission Type:", choices = unique(cars_data$Transmission)),
      selectInput("color", "Select Car Color:", choices = unique(cars_data$Color)),
      selectInput("Body.Style", "Select Car Body Style:", choices = unique(cars_data$Body.Style))),
    mainPanel(
      textOutput("price_output")
    )
  )
)

server_his_price <- function(input, output) {
  output$price_output <- renderText({
    filtered_data <- cars_data %>%
      filter(Dealer_Name == input$dealer_name,
             Company == input$company,
             Model == input$model,
             Transmission == input$transmission,
             Color == input$color,
             Body.Style == input$Body.Style)
    
    if (nrow(filtered_data) > 0) {
      paste("The price is:", paste(unique(filtered_data$Price), collapse = ", "), "USD")
    } else {
      "No matching car price information"
    }
  })
}

shinyApp(ui = ui_his_price, server = server_his_price)

```

```{r}
# Distribution of sale by date
ui_date <- fluidPage(
  titlePanel("Car Sales Dashboard"),
  sidebarLayout(
    sidebarPanel(
      selectInput("company", "Select Company:", choices = unique(cars_data$Company)),
      sliderInput("price_range", "Select Price Range:", min = 0, max = 500000, value = c(0, 50000))
    ),
    mainPanel(
      plotOutput("sales_plot")
    )
  )
)


server_date <- function(input, output) {

  filtered_data <- reactive({
    filter(cars_data, Company == input$company & Price >= input$price_range[1] & Price <= input$price_range[2])
  })
  

  output$sales_plot <- renderPlot({
    ggplot(filtered_data(), aes(x = Date, y = Price, color = Dealer_Name)) +
      geom_point() +
      labs(title = "Sales by Date", x = "Date", y = "Price")
  })
}


shinyApp(ui = ui_date, server = server_date)

```
# Data visualization 
Car sales are typically slow at the beginning of the year, peak in September, decline in October, and then reach another peak in November and December, winter is the busiest season of the year.  The top three best-selling car brands are Chevrolet, Dodge, and Ford. The top 10 selling car models are Diamante, Prizm , Silhouette, Passat, Ram Pickup, Jetta ,RL ,LS400, LHS, A6,respectively. The region with the highest sales revenue is Austin, Janesville, Scottsdale. The majority of car buyers are male, nearly four times greater than female buyers. Car price around 20000 seems most popular. The most popular color is pale white among both female and male. Then black and red. People more preferred auto transmission. The best seller car body style is SUV, then hatchback.

```{r}
# Aggregate sales count by month
cars_data$month <- month(cars_data$Date)
monthly_sales <- cars_data %>%
  group_by(month) %>% 
  summarize(Total_Count = n())

# Plotting sales count by month to show seasonal patterns
ggplot(monthly_sales, aes(x = month, y = Total_Count)) +
  geom_line() +
  labs(title = 'Monthly Sales Patterns', x = 'Month', y = 'Total Count') +
  theme_minimal() +
  scale_x_continuous(breaks = 1:12)
```

```{r}
# Distribution of Car Company
company_counts <- cars_data %>%
  count(Company) %>%
  arrange(desc(n))

ggplot(cars_data, aes(x = factor(Company, levels = company_counts$Company), fill = factor(Company))) +
  geom_bar() +
  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5, size = 2, color = 'black')+
  labs(title = "Distribution of Car Company", x = "Company", y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
The top three best-selling car brands are Chevrolet, Dodge, and Ford.

```{r}
# Top 10 Selling Car Models
top_models <- cars_data %>%
  count(Model) %>%
  arrange(desc(n)) %>%
  head(10)

ggplot(top_models, aes(x = reorder(Model, n), y = n, fill = Model)) +
  geom_col() +
  geom_text(aes(label = n), vjust = -0.5, size = 3, color = "black") +
  labs(title = "Top 10 Selling Car Models", x = "Model", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +  
  coord_flip()
```
The top 10 selling car models are Diamante, Prizm , Silhouette, Passat, Ram Pickup, Jetta ,RL ,LS400, LHS, A6,respectively.

```{r}
# Distribution of Dealer Region
ggplot(cars_data, aes(x = Dealer_Region, fill = Dealer_Region)) +
  geom_bar() +
  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5, size = 3, color = 'black')+
  labs(title = "Distribution of Car Dealer Region", x = "Dealer Region", y = "Count") +
  theme_bw()
```
The region with the highest sales revenue is Austin, Janesville, Scottsdale.

```{r}
# Distribution of Gender among car buyers
gender_counts <- table(cars_data$Gender)

pie(gender_counts, labels = paste(names(gender_counts), "(", round(gender_counts/sum(gender_counts)*100, 1), "%)", sep = ""), col = c("pink", "lightblue"))
title("Distribution of Gender among car buyers")
legend("topright", legend = c("Female", "Male"), fill = c("pink", "lightblue"))

```
The majority of car buyers are male, nearly four times more than female buyers.

```{r}
# Car Price Distribution
ggplot(cars_data, aes(x = Price)) +
  geom_histogram(binwidth = 5000, fill = "skyblue", color = "black") +
  labs(title = "Car Price Distribution", x = "Price", y = "Count") +
  theme_minimal()
```
Car price around 20000 seems most popular.

```{r}
# Distribution of Annual Income
incomex <- cars_data$Annual.Income/10000000
ggplot(cars_data, aes(x = `incomex`, y = "")) +
  geom_boxplot() +
  labs(title = "Distribution of Annual Income", x = "Annual Income") +
  theme(axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank())
```

```{r}
# Distribution of Annual Income
ggplot(cars_data, aes(x = `incomex`)) +
  geom_histogram(bins = 70, fill = "skyblue", color = "black") +
  labs(title = "Distribution of Annual Income", x = "Annual Income", y = "Count") 

```

```{r}
# Distribution of Car color and Proportion of Gender
gender_color <- cars_data %>%
group_by(Gender, Color) %>%
summarize(count = n())

ggplot(gender_color, aes(x = Color, y = count, fill = factor(Gender))) +
geom_bar(stat ="identity", position = "stack") +
labs(title = "Distribution of Car color and Proportion of Gender", x = "Color", y = "Count")
```
The most popular color is pale white among both female and male. Then black and red.
```{r}
# Distribution of Car Transmission
Transmission_counts <- cars_data %>%
  group_by(Transmission) %>%
  summarize(count = n())

ggplot(Transmission_counts, aes(x = "", y = count, fill = Transmission)) +
  geom_bar(stat = "identity", width = 1) +
  labs(title = "Distribution of Car Transmission", fill = "Transmission") +
  coord_polar(theta = "y") +
  theme_void() +
  geom_text(aes(label = count), position = position_stack(vjust = 0.5))
```
People more prefered auto transmission.
```{r}
# Distribution of Car Body Style
ggplot(cars_data, aes(x = Body.Style, fill = Body.Style)) +
  geom_bar() +
  labs(title = "Distribution of Car Body Style", x = "Car Body Style", y = "Count") +
  theme_bw()
```
The best seller car body style is SUV, then hatchback.

# Data preprocessing
For data preprocessing, we converted Date column to date format assuming consistently format, modified the Engine column which contains special characters, and encoded some string columns. In final feature selection, we delete car id, data, customer name, engine, dealer number and phone columns, and the rest columns will be the feature colunm.
```{r}
cars_data <- cars_data %>%
  mutate(
    Gender = as.numeric(factor(Gender, levels = c("Male", "Female"), labels = c(1, 0))),
    Dealer_Name = as.numeric(factor(Dealer_Name)),
    Company = as.numeric(factor(Company)),
    Model = as.numeric(factor(Model)),
    Transmission = as.numeric(factor(Transmission)),
    Color = as.numeric(factor(Color)),
    Body.Style = as.numeric(factor(Body.Style)),
    Dealer_Region = as.numeric(factor(Dealer_Region))
  )
```



# Methodology

For model building, we used random forest regression try to predict the car price. From original dataset, we can notice that the car price is always a whole number. However, the model predict car price always include decimal point. So we considered when predict price minus original price is less than 500, then will considered the predict price is consistent with original price. So the final accuracy is 94.66%. This model can be used to predict car price and also future market trends.

```{r}
library(randomForest)
set.seed(123) 
filter_data <- select(cars_data, -Car_id, -Date, -Customer.Name, -Engine, -Dealer_No, -Phone)
train_index <- sample(1:nrow(filter_data), 0.7 * nrow(filter_data)) 
train_data <- filter_data[train_index, ]
test_data <- filter_data[-train_index, ]
dim(train_data)
dim(test_data)
```

```{r}
rf_model <- randomForest(Price ~ ., data = train_data)

predictions <- predict(rf_model, newdata = test_data)

threshold <- 500
accuracy <- mean(abs(predictions - test_data$Price) <= threshold)
print(paste("Accuracy (within $500):", accuracy))
```

```{r}
# Car price prediction among different dealer
ui_predict <- fluidPage(
  titlePanel("Car Price Prediction"),
  sidebarLayout(
    sidebarPanel(
      selectInput("dealer", "Dealer Name", choices = unique(cars_data$Dealer_Name)),
      selectInput("company", "Company", choices = unique(cars_data$Company)),
      selectInput("model", "Model", choices = unique(cars_data$Model)),
      selectInput("transmission", "Transmission", choices = unique(cars_data$Transmission)),
      selectInput("body_style", "Body Style", choices = unique(cars_data$Body.Style)),
      actionButton("predict_button", "Predict")
    ),
    mainPanel(
      textOutput("predicted_price")
    )
  )
)

server_predict <- function(input, output) {

  filtered_data <- reactive({
    filter(data, Dealer_Name == input$dealer &
                    Company == input$company &
                    Model == input$model &
                    Transmission == input$transmission &
                    Body.Style == input$body_style)
  })
  
  # Predict price using random forest model
  output$prediction_text <- renderText({
    rf_model <- randomForest(Price ~ ., data = cars_data)
    prediction <- predict(rf_model, newdata = filtered_data())
    paste("Predicted Price:", prediction, "USD")
  })
}

shinyApp(ui = ui_predict, server = server_predict)


```





